---
title: "Perceived sensitivity of brain data"
author: "Shenyang Huang"
date: "09/04/2023"
output:
  html_document: default
  word_document: default
  pdf_document: default
---
Script written 07/20/2021 for first submission
Script updated 09/04/2023 for JLB revision 1

# Set up

## - R library
```{r setup, message=FALSE, warning=FALSE}
# tidy data
library(MASS)
library(tidyr)
library(tidyverse)

# factor analysis
library(paran)
library(psych)
library(GPArotation)

# mixed model
library(lme4)
library(lmerTest)
library(emmeans)
library(car)

# figures
library(ggeffects)
library(ggplot2)
library(ggpubr)
library(gghalves)
library(cowplot)
library(patchwork)
rm(list=ls())
invisible(gc())

knitr::opts_chunk$set(echo = F)

```


## - load full data to exclude participants
```{r}
df_sens_full <- read.csv("data/Final Brain Data Sensitivity for SSI_October 19, 2020_15.16 (4).csv") %>%
  mutate(Duration..in.seconds. = strtoi(Duration..in.seconds.),
         Progress = strtoi(Progress))

### remove two extra "header rows"
df_sens_selected <- df_sens_full[3:nrow(df_sens_full),] 
n_recruited <- nrow(df_sens_selected) 

### 1. remove those who did not consent
# consented <- !grepl("DO NOT", df_sens$consent)
consented <- grepl("I agree to", df_sens_full$consent)
df_sens_selected <- df_sens_full[consented, ] 

### 2. remove those that did not finish
finished <- df_sens_selected$Finished=="True"
df_sens_selected <- df_sens_selected[finished, ] 

### 3. failed manipulation check
passed_check <- df_sens_selected$sensitive.informatio_36=="somewhat sensitive\n"
df_sens_selected <- df_sens_selected[passed_check, ]

### 4. completion time
too_slow <- df_sens_selected$Duration..in.seconds. > 31*60
too_fast <- df_sens_selected$Duration..in.seconds. < 3*60
df_sens_selected <- df_sens_selected[!(too_slow | too_fast), ]

cat(
  sprintf("Recruited: %s \n", n_recruited),
  sprintf("Did not consent: %s \n", sum(!consented)),
  sprintf("Did not finish: %s \n", sum(!finished)),
  sprintf("Failed attention check: %s \n", sum(!passed_check)),
  sprintf("> 31 min: %s \n", sum(too_slow)),
  sprintf("< 3 min: %s \n", sum(too_fast)),
  sprintf("excluded: %s \n",
          sum(!consented)+sum(!finished)+sum(!passed_check)+sum(too_slow)+sum(too_fast)),
  sprintf("remaining: %s \n", nrow(df_sens_selected))
)
# View(df_sens)

```


## - format the dataframe
```{r}
df_sens <- df_sens_selected %>%
  select(
    pid,
    ### demographics
    Gender, Race, Age, Edu, Income, marital, PO,
    ### perception of sensitive information: 35 items + 1 attention check
    starts_with("sensitive.informatio_"), -sensitive.informatio_36,
    ### other measures pertaining to tech
    EarAdopt, # whether one consider themselves an early adoptor of tech
    WearTech, # do they use wearable tech
  ) %>%
  rename(ID = pid,
         Marital = marital) %>%
  ### use the midpoint of the levels and transform age to a continous variable
  mutate(Age_rescale = case_when(
    Age=="18-24" ~ 21,
    Age=="25-34" ~ 29.5,
    Age=="35-44" ~ 39.5,
    Age=="45-54" ~ 49.5,
    Age=="55-64" ~ 59.5,
    Age=="65+" ~ 70)) %>%
  ### convert to ordinal variable 1-6, six age groups
  mutate(AgeLvl = case_when(
    Age=="18-24" ~ 1,
    Age=="25-34" ~ 2,
    Age=="35-44" ~ 3,
    Age=="45-54" ~ 4,
    Age=="55-64" ~ 5,
    Age=="65+" ~ 6)) %>%
  ### five levels to three
  mutate(EarAdoptLvl = case_when(
    grepl("generally the first", EarAdopt) ~ 1,
    grepl("among the first", EarAdopt) ~ 2,
    grepl("in the middle", EarAdopt) ~ 3,
    grepl("among the last", EarAdopt) ~ 4,
    grepl("generally the last", EarAdopt) ~ 5)) %>%
  ### binary categorization based on "college degree" or higher
  mutate(CollegeDegree = case_when(
    Edu %in% c("Graduate/Professional school or beyond ",
               "College graduate") ~ "college degree",
    Edu %in% c("Less than high school",
               "High school graduate or equivalent",
               "Some college ") ~ "no college degree")) %>%
  ### order income 1-9, nine groups
  mutate(IncomeLvl = case_when(
    Income=="Less than $15,000 " ~ 1,
    Income=="$15,000 - $24,999 " ~ 2,
    Income=="$25,000 - $34,999 " ~ 3,
    Income=="$35,000 - $49,999 " ~ 4,
    Income=="$50,000 - $74,999 " ~ 5,
    Income=="$75,000 - $99,999 " ~ 6,
    Income=="$100,000 - $149,999 " ~ 7,
    Income=="$150,000 - $199,999" ~ 8,
    Income=="More than $200,000" ~ 9)) %>%
  ### transform the sensitive information columns from textual to numeric
  mutate(across(starts_with("sensitive.informatio_"), 
                \(x) case_when(
                  grepl("very sensitive", x) ~ 4,
                  grepl("somewhat sensitive", x) ~ 3,
                  grepl("not too sensitive", x) ~ 2,
                  grepl("not at all sensitive", x) ~ 1
                ))) %>%
  ### rearrange columns
  relocate(Age_rescale, AgeLvl, .after=Age) %>%
  relocate(EarAdoptLvl, .after=EarAdopt) %>%
  relocate(CollegeDegree, .after=Edu) %>%
  relocate(IncomeLvl, .after=Income)


### rename sensitive information 1-35 based on the actual item
# View(df_sens)


### use the two extra "header rows" to match sensitive information 1-36 to actual items
df_sens_item <- df_sens_full[1, ] %>% 
  select(starts_with("sensitive.informatio_")) %>%
  mutate(across(everything(), \(x) str_split(x, "have access to it. - ", simplify = TRUE)[,2])) %>%
  t() %>% 
  data.frame() %>% 
  rownames_to_column(var="SI") %>%
  rename("Item"=X1)

### convert column names from 1-36 to actual items
### and use more readable labels for each SI question
SI_labels <- list(
  "_1" = c("SI_1_SSN", "Social Security Number"),
  "_2" = c("SI_2_health", "Health"),
  "_4" = c("SI_4_phone", "Phone Conversations"), 
  "_5" = c("SI_5_email", "Email Messages"),
  "_6" = c("SI_6_location", "Location"),
  "_7" = c("SI_7_texts", "Text Messages"),
  "_8" = c("SI_8_numbers", "Phone Numbers Called/Texted"),
  "_9" = c("SI_9_birthdate", "Birth Date"),
  "_10" = c("SI_10_relationship", "Relationship History"),
  "_11" = c("SI_11_website", "Websites Visited"),
  "_12" = c("SI_12_search", "Online Search History"),
  "_13" = c("SI_13_religion", "Religious Views"),
  "_14" = c("SI_14_friends", "Friends"),
  "_15" = c("SI_15_politics", "Political Views"),
  "_16" = c("SI_16_media", "Media Preferences"),
  "_17" = c("SI_17_purchases", "Purchasing Habits"),
  "_18" = c("SI_18_credit", "Credit Score"),
  "_19" = c("SI_19_drowsiness", "Drowsiness"),
  "_20" = c("SI_20_alertness", "Alertness"),
  "_21" = c("SI_21_focus", "Focus"),
  "_22" = c("SI_22_concentration", "Concentration"),
  "_24" = c("SI_24_brainwave", "Brainwaves"),
  "_25" = c("SI_25_brain.health", "Brain Health"),
  "_26" = c("SI_26_substance", "Substance Use"),
  "_27" = c("SI_27_emotion", "Emotions"),
  "_28" = c("SI_28_anxiety", "Anxiety"),
  "_29" = c("SI_29_mental.image", "Images in mind"),
  "_30" = c("SI_30_mental.thoughts", "Thoughts in mind"),
  "_31" = c("SI_31_intelligence", "Intelligence Quotient"),
  "_32" = c("SI_32_sleep", "Sleep Patterns"),
  "_33" = c("SI_33_sexual.orientation", "Sexual Orientation"),
  "_34" = c("SI_34_genetic", "Genetic information"),
  "_35" = c("SI_35_social.media", "Social Media Contents")
)

### renaming
for (i in 1:length(SI_labels)) {
  item <- names(SI_labels)[i]
  original_colname <- paste0("sensitive.informatio", item)
  new_colname <- SI_labels[[item]][1]
  df_sens[new_colname] <- df_sens[original_colname]
}
df_sens <- df_sens %>% 
  select(-starts_with("sensitive.informatio_"))

### save data on disk
write.csv(df_sens, "data/data_sensitivity.csv", row.names = F)
rm(list=ls()[!grepl('SI_labels', ls())])

```


## - load data and format
```{r}
df_sens <- read.csv("data/data_sensitivity.csv") %>%
  mutate(ID = 
           factor(ID),
         WearTech = 
           factor(WearTech,
                  levels=c("No", "Yes")),
         CollegeDegree = 
           factor(CollegeDegree,
                  levels=c("no college degree", "college degree")))

```


## - demographics
```{r, echo=FALSE}
cat(sprintf(
  "Gender breakdown: %i Female, %i Male, and %i Other. 
  \n All of the following demographic breakdown will be in the percentage of the total sample size. 
The age breakdown of our participants was as follows: 18 – 24 = %.1f%%, 25 – 34 = %.1f%%, 35 – 44 = %.1f%%, 45 – 54 = %.1f%%, 55 – 64 = %.1f%%, and 65+ = %.1f%%. 
The racial breakdown of our participants is as follows: White = %.1f%%, Black = %.1f%%, Hispanic/Latino = %.1f%%, Asian = %.1f%%, American Indian/Alaska Native = %.1f%%, Native Hawaiian/other Pacific Islander = %.1f%%, and Other = %.1f%%. 
The educational breakdown of our participants is as follows: less than high school = %.1f%%, High school graduate or equivalent = %.1f%%, Some college = %.1f%%, College graduate = %.1f%%, and Graduate/Professional school or beyond = %.1f%%.
\n 
The household income breakdown in percentage was as follows: Less than $15,000 = %.1f%%, $15,000 – $24,999 = %.1f%%, $25,000 – $34,999 = %.1f%%, $35,000 – $49,999 = %.1f%%, $50,000 – $74,999 = %.1f%%, $75,000 – $99,999 = %.1f%%, $100,000 – $149,999 = %.1f%%, $150,000 – $199,999 = %.1f%%, and %.1f%% of participants make more than $200,000. 
The marital status breakdown of the participants was as follows: Never Married = %.1f%%, Married = %.1f%%, Separated = %.1f%%, Divorced = %.1f%%, and Widowed = %.1f%%. Lastly, the political affiliation breakdown of our participants is as follows: Democrat = %.1f%%, Independent = %.1f%%, Republican = %.1f%%, and Other = %.1f%%. 
Table X below presents the demographic breakdown of the participants within this study by gender.",
sum(df_sens$Gender=="Female"),
sum(df_sens$Gender=="Male"),
sum(df_sens$Gender=="Other"),
mean(df_sens$Age=="18-24")*100,
mean(df_sens$Age=="25-34")*100,
mean(df_sens$Age=="35-44")*100,
mean(df_sens$Age=="45-54")*100,
mean(df_sens$Age=="55-64")*100,
mean(df_sens$Age=="65+")*100,
mean(df_sens$Race=="White")*100,
mean(df_sens$Race=="Black ")*100,
mean(df_sens$Race=="Hispanic or Latino")*100,
mean(df_sens$Race=="Asian ")*100,
mean(df_sens$Race=="American Indian or Alaska Native ")*100,
mean(df_sens$Race=="Native Hawaiian or Other Pacific Islander ")*100,
mean(df_sens$Race=="Other ")*100,
mean(df_sens$Edu=="Less than high school")*100,
mean(df_sens$Edu=="High school graduate or equivalent")*100,
mean(df_sens$Edu=="some college")*100,
mean(df_sens$Edu=="College graduate")*100,
mean(df_sens$Edu=="Graduate/Professional school or beyond ")*100,
mean(df_sens$Income=="Less than $15,000 ")*100,
mean(df_sens$Income=="$15,000 - $24,999 ")*100,
mean(df_sens$Income=="$25,000 - $34,999 ")*100,
mean(df_sens$Income=="$35,000 - $49,999 ")*100,
mean(df_sens$Income=="$50,000 - $74,999 ")*100,
mean(df_sens$Income=="$75,000 - $99,999 ")*100,
mean(df_sens$Income=="$100,000 - $149,999 ")*100,
mean(df_sens$Income=="$150,000 - $199,999")*100,
mean(df_sens$Income=="More than $200,000")*100,
mean(df_sens$Marital=="Never married")*100,
mean(df_sens$Marital=="Married")*100,
mean(df_sens$Marital=="Separated")*100,
mean(df_sens$Marital=="Divorced")*100,
mean(df_sens$Marital=="Widowed")*100,
mean(df_sens$PO=="Democrat")*100,
mean(df_sens$PO=="Independent")*100,
mean(df_sens$PO=="Republican")*100,
mean(df_sens$PO=="Other")*100
))

```



# Analyses

# Individual Sensitivity questions

## - full labels
```{r}
SI_items <- colnames(df_sens)
SI_items <- SI_items[grepl("SI_", SI_items)]

SI_labels_full <- c()
for (i in 1:length(SI_labels)) {
  SI_labels_full <- c(SI_labels_full, SI_labels[[i]][2])
}
names(SI_labels_full) <- SI_items
SI_labels_full

```


## - percentage of responses, mean, and sd
```{r}
df_plot_SI_Qns_proportion <- data.frame(
  cbind(1:4,
        c("not at all sensitive",
          "not too sensitive",
          "somewhat sensitive",
          "very sensitive"),
        apply(df_sens[SI_items], 2, table)/nrow(df_sens))) %>%
  rename(Sensitivity_num=V1,
         Sensitivity_text=V2) %>%
  pivot_longer(starts_with("SI_"),
               names_to = "Item",
               values_to = "proportion") %>%
  mutate(Sensitivity_num = as.numeric(Sensitivity_num),
         proportion = as.numeric(proportion))
# View(df_plot_SI_Qns_percent)

### compute mean, sd, and sem
df_plot_SI_Qns_meansd <- data.frame(
  cbind(apply(df_sens[SI_items], 2, mean),
        apply(df_sens[SI_items], 2, sd))) %>%
  rownames_to_column(var="Item") %>%
  rename(mean=X1, sd=X2) %>%
  mutate(se  = round(sd/sqrt(nrow(df_sens)-1), 2),
         mean = round(mean, 2),
         sd = round(sd, 2)) %>%
  mutate(Description = SI_labels_full[Item]) %>%
  relocate(Description, .after=Item) 

### combine both dataframes
df_plot_SI_Qns <- left_join(
  df_plot_SI_Qns_proportion,
  df_plot_SI_Qns_meansd
) %>%
  arrange(-mean) 

### order items by highest to lowest proportion of 4s
Item_order <- df_plot_SI_Qns %>% 
  filter(Sensitivity_num==4) %>%
  arrange(proportion) %>%
  select(Item) %>%
  unlist() %>% 
  as.character()

df_plot_SI_Qns <- df_plot_SI_Qns %>%
  mutate(Item=factor(Item, levels=Item_order))

write.csv(df_plot_SI_Qns, 'data/Figure1_pct_mean_sd.csv', row.names=F)
df_plot_SI_Qns

```


## - plot by percentage 
```{r}
### arrange items by their proportion of "very sensitive" responses
g_SI_pct <- df_plot_SI_Qns %>%
  ggplot(aes(x=Item, 
             y=proportion, 
             fill=Sensitivity_text)) +
  geom_bar(stat="identity") +
  scale_x_discrete("Information Items Presented",
                   labels=SI_labels_full) +
  scale_y_continuous("Percentage of Responses",
                     labels = scales::percent) +
  scale_fill_discrete("Sensitivity Ratings") +
  scale_fill_brewer(NULL,
                    palette = 7,
                    breaks=c(
                      "very sensitive",
                      "somewhat sensitive",
                      "not too sensitive",
                      "not at all sensitive")) +
  ggtitle("Sensitivity Ratings of All Information Questions") +
  coord_flip(expand = FALSE) +
  theme_pubclean() +
  theme(legend.position="bottom",
        axis.ticks = element_blank(),
        plot.margin = margin(0, 20, 0, 0)) +  
  NULL

g_SI_pct
ggsave("figures/SI_pct.tiff", g_SI_pct, dpi=600, width=10, height=7)

```



# Exploratory Factor Analysis 

## - EFA on 33 items and plot factor loadings
```{r}
### identifies latent variables supposedly exert causal influence on observed variables and influence covariance 

### determine the number of factors to retain
n_factors <- paran(df_sens[SI_items], seed=1, graph=TRUE)
n_factors$Retained # 3 factors retained

### run EFA
theta <- 0.4 # threshold for loading to be considered
fa_result <- fa(df_sens[SI_items], 
                n_factors$Retained, 
                rotate="promax", 
                scores="Bartlett")
df_fa_result <- data.frame(fa_result$loadings[]) %>%
  rownames_to_column(var="Item") %>%
  rename(F1=MR1,
         F2=MR2,
         F3=MR3) %>% 
  mutate(pass1 = cut(F1, breaks = c(-Inf, theta, Inf),
                     labels = c("low", "high"),
                     right = TRUE),
         pass2 = cut(F2, breaks = c(-Inf, theta, Inf),
                     labels = c("low", "high"),
                     right = TRUE),
         pass3 = cut(F3, breaks = c(-Inf, theta, Inf),
                     labels = c("low", "high"),
                     right = TRUE))

### reorder load_df into chunks of three factors
df_fa1 <- df_fa_result %>% filter(pass1=="high") %>% 
  arrange(-F1)
df_fa2 <- df_fa_result %>% filter(pass2=="high") %>% 
  arrange(-F2)
df_fa3 <- df_fa_result %>% filter(pass3=="high") %>% 
  arrange(-F3)
df_fa0 <- df_fa_result %>% filter(pass1=="low", 
                                  pass2=="low", 
                                  pass3=="low") %>%
  arrange(-F3)


### In the figure, plot Public F3, Private F2, and Brain F1
# df_fa_reordered <- rbind(df_fa3, df_fa2, df_fa1, df_fa0)
df_fa_reordered <- rbind(df_fa1, df_fa2, df_fa3, df_fa0)

for (i in 1:3) {
  g <- ggbarplot(data=df_fa_reordered,
                 x="Item",
                 y=paste0("F", i),
                 fill = paste0("pass", i),
                 color = NULL) +
    coord_flip() +
    scale_x_discrete(labels=SI_labels_full) +
    scale_y_continuous(name = "Loading", 
                       limits = c(-0.25, 0.86),
                       breaks = c(-0.2, 0, 0.4, 0.8)) +
    theme_get() +
    theme(plot.title = element_text(hjust=0.38)) +
    rremove("legend") +
    rremove("ylab") +
    NULL
  assign(paste0("g_F", i), g) 
}
g_F3 <- g_F3 + 
  scale_fill_manual(values=ggplot2::alpha(c("gray85", "#00BA38"), 0.7)) +
  ggtitle("Public") + 
  rremove("xlab")
g_F2 <- g_F2 + 
  scale_fill_manual(values=ggplot2::alpha(c("gray85", "#619CFF"), 0.7)) +
  ggtitle("Private") + 
  rremove("y.text") + 
  rremove("y.ticks") 
g_F1 <- g_F1 + 
  scale_fill_manual(values=ggplot2::alpha(c("gray85", "#F8766D"), 0.7)) +
  ggtitle("Mental/Affective") + 
  rremove("y.text") + 
  rremove("y.ticks") + 
  rremove("xlab")

(g_F3 | g_F2| g_F1)
ggsave("figures/SI_factor_loading.tiff", dpi=600, width=8, height=5)


```


## - factor score using weighted average
```{r}
# REF: https://scholarworks.umass.edu/cgi/viewcontent.cgi?article=1226&context=pare
loadings_all <- fa_result$loadings
load1_passed <- abs(loadings_all[,1]) >= theta ### mental/affective
load2_passed <- abs(loadings_all[,2]) >= theta ### public
load3_passed <- abs(loadings_all[,3]) >= theta ### private

df_sens_fa <- df_sens %>% mutate(
  ### Bartlett method
  SI_factor1 = fa_result$scores[,1],
  SI_factor2 = fa_result$scores[,2],
  SI_factor3 = fa_result$scores[,3],
  
  ### weighted average based on factor loadings (if > 0.4)
  SI_factor1_wa = 
    rowSums(loadings_all[,1][load1_passed] * df_sens[SI_items][load1_passed]) / sum(loadings_all[,1][load1_passed]),
  
  SI_factor2_wa = 
    rowSums(loadings_all[,2][load2_passed] * df_sens[SI_items][load2_passed]) / sum(loadings_all[,2][load2_passed]),
  
  SI_factor3_wa = 
    rowSums(loadings_all[,3][load3_passed] * df_sens[SI_items][load3_passed]) / sum(loadings_all[,3][load3_passed]))

### high correlation between Bartlett and weighted sum methods
### however, weighted averages allow for cross-factor comparisons
cor(df_sens_fa[c("SI_factor1", "SI_factor1_wa")]) %>% round(3)
cor(df_sens_fa[c("SI_factor2", "SI_factor2_wa")]) %>% round(3)
cor(df_sens_fa[c("SI_factor3", "SI_factor3_wa")]) %>% round(3)


```


## - long format data for regression
```{r}
df_sens_fa_long <- df_sens_fa %>%
  select(-starts_with("SI_"), 
         SI_factor1_wa, SI_factor2_wa, SI_factor3_wa) %>%
  pivot_longer(cols=starts_with("SI_factor"),
               names_pattern = "^SI_(.*)_wa", 
               names_to = "Factor",
               values_to = "Factor_score_wa") %>%
  mutate(Factor = factor(Factor,
                         levels=c("factor1", "factor2", "factor3")),
         ### re-order the three factors
         Factor_num = case_when(
           Factor=="factor2" ~ 1,
           Factor=="factor3" ~ 2,
           Factor=="factor1" ~ 3),
         Factor_name = case_when(
           Factor=="factor2" ~ "Public",
           Factor=="factor3" ~ "Private",
           Factor=="factor1" ~ "Mental/Affective"))

```


## - plot factor scores
```{r}
### compute error bars
errorbars <- df_sens_fa_long %>%
  group_by(Factor_name, Factor_num) %>%
  summarise(n=length(Factor_score_wa),
            mean=mean(Factor_score_wa),
            sd=sd(Factor_score_wa),
            se=sd/sqrt(n-1),
            lower=mean-sd,
            upper=mean+sd) %>%
  ungroup()

### make plot
g_factors_wa <- ggplot(df_sens_fa_long, 
                       aes(x=Factor_num, 
                           y=Factor_score_wa,
                           fill=factor(Factor_num), 
                           color=factor(Factor_num))) +
  geom_point(position=position_jitter(width=.1), size=0.4, alpha=0.4) +
  geom_half_boxplot(aes(x=Factor_num+0.2),
                    position=position_nudge(x=-0.05), 
                    side="r", outlier.shape = NA, center=TRUE, 
                    errorbar.draw = FALSE, width=.1, alpha=0.4) +
  geom_half_violin(position=position_nudge(x=.25), side="r", alpha=0.4) +
  geom_point(data=errorbars, 
             aes(x=Factor_num, y=mean), 
             position = position_nudge(x = 0.3), size = 2.5) +
  geom_errorbar(data=errorbars, 
                aes(x=Factor_num,
                    y=mean,
                    ymin=lower, 
                    ymax=upper),
                position = position_nudge(x = 0.3), width = 0) +
  scale_x_continuous(name="Type of Information",
                     breaks=1:3,
                     labels=c(
                       "1"="Public",
                       "2"="Private",
                       "3"="Mental/Affective")) +
  scale_y_continuous(name="", 
                     limits = c(0.9, 4.1),
                     breaks = 1:4,
                     labels = c("1=\nnot at all sensitive",
                                "2=\nnot too sensitive",
                                "3=\nsomewhat sensitive",
                                "4=\nvery sensitive")) +
  scale_fill_manual(values=c("1"="#00BA38",
                             "2"="#619CFF",
                             "3"="#F8766D")) +
  scale_color_manual(values=c("1"="#00BA38",
                              "2"="#619CFF",
                              "3"="#F8766D")) +
  ggtitle("Perceived sensitivity of different types of information") +
  theme_cowplot() +
  theme(axis.text.y = element_text(hjust = 0.5, size=10)) +
  guides(fill="none", color="none") +
  NULL

g_factors_wa
ggsave("figures/SI_factors_distribution.tiff", g_factors_wa,
       dpi=600, width=8, height=5)

```



# Statistical analyses using factor scores

## - compare three factor scores
```{r}
### use factor1 (mental/affective) as reference
df_sens_fa_long$Factor_name <- relevel(factor(df_sens_fa_long$Factor_name),
                                       ref="Mental/Affective")

### use the default treatment coding
contrasts(df_sens_fa_long$Factor_name)

### alternatively one could first check the main effect of Factor across three levels and then use emmeans to get pairwise comparisons; but it doesn"t really matter
m_fs <- lmer(Factor_score_wa ~ Factor_name + (1|ID), data=df_sens_fa_long)

Anova(m_fs, type=3)

round(coef(summary(m_fs)), 4)

### 95% confidence intervals around the estimates
confint(m_fs, "beta_", level=0.95, 
        method="boot", nsim=1000, boot.type="perc", 
        FUN=NULL, quiet=FALSE, oldNames=TRUE) %>% 
  round(2)


```


## - compare factors by SES, focusing on factor 1 (reference level) 
```{r}
### ignore Other-Gender since there are only 5 such subjects--not enough data
df_sens_fa_long <- df_sens_fa_long %>%
  mutate(
    Gender_MF = relevel(factor(ifelse(Gender=="Other", NA, Gender)), 
                        ref="Female"),
    CollegeDegree = relevel(factor(CollegeDegree), 
                            ref="college degree"))

### use deviation coding of categorical variables
### female = 0.5, male = -0.5
contrasts(df_sens_fa_long$Gender_MF) <- contr.sum(2)/2 
### college degree = 0.5, no cd = -0.5
contrasts(df_sens_fa_long$CollegeDegree) <- contr.sum(2)/2

```


## - determine whether to include the interaction between SES variables or not
```{r}
### only interaction between Factor and SES
m_fs_by_SES <- lmer(Factor_score_wa ~ 
                      Factor_name + 
                      scale(AgeLvl) + 
                      scale(IncomeLvl) + 
                      CollegeDegree + 
                      Gender_MF +
                      Factor_name:scale(AgeLvl) + 
                      Factor_name:scale(IncomeLvl) + 
                      Factor_name:CollegeDegree + 
                      Factor_name:Gender_MF +
                      (1|ID),
                    data=df_sens_fa_long)

### two-way interactions of SES as well
m_fs_by_SES_2 <- lmer(Factor_score_wa ~ 
                        Factor_name + 
                        scale(AgeLvl) + 
                        scale(IncomeLvl) + 
                        CollegeDegree + 
                        Gender_MF +
                        Factor_name:scale(AgeLvl) + 
                        Factor_name:scale(IncomeLvl) + 
                        Factor_name:CollegeDegree +
                        Factor_name:Gender_MF +
                        scale(AgeLvl):scale(IncomeLvl) + 
                        scale(AgeLvl):CollegeDegree + 
                        scale(AgeLvl):Gender_MF +
                        scale(IncomeLvl):CollegeDegree + 
                        scale(IncomeLvl):Gender_MF +
                        CollegeDegree:Gender_MF + 
                        (1|ID),
                      data=df_sens_fa_long)

### check the impact of including the 2-way SES interactions -- no 
anova(m_fs_by_SES, m_fs_by_SES_2)
### no significant difference, thus consider the simpler model 

round(coef(summary(m_fs_by_SES)), 4)

confint(m_fs_by_SES, "beta_", level=0.95,
        parm = c("(Intercept)", "scale(AgeLvl)", 
                 "scale(IncomeLvl)", "CollegeDegree1", "Gender_MF1"),
        method="boot", nsim=1000, boot.type="perc",
        FUN=NULL, quiet=FALSE, oldNames=TRUE) %>%
  round(2)

```


## - plot Factor 1 "Mental/Affective" by SES - lmer 
```{r}
### get marginal means
df_plot_m_fs_by_SES <- ggemmeans(m_fs_by_SES, 
                                 ~ AgeLvl + CollegeDegree + Gender_MF + Factor_name) %>% 
  data.frame() %>% 
  rename(CollegeDegree=group,
         Gender=facet) %>%
  filter(panel=="Mental/Affective") # only look at factor 1

### plot
df_plot_m_fs_by_SES %>%
  ggplot(aes(x=x, y=predicted, fill=CollegeDegree, color=CollegeDegree)) +
  facet_wrap(~Gender, nrow = 1) +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high, color=NULL), alpha=0.3) +
  geom_line() +
  scale_x_continuous(name="Age range",
                     breaks=1:6,
                     labels=c("1"="18-24",
                              "2"="25-34",
                              "3"="35-44",
                              "4"="45-54",
                              "5"="55-64",
                              "6"="65+")) +
  scale_y_continuous(name="Perceived sensitivity", 
                     limits = c(1, 4),
                     breaks = 1:4,
                     labels = c("1=\nnot at all sensitive",
                                "2=\nnot too sensitive",
                                "3=\nsomewhat sensitive",
                                "4=\nvery sensitive")) +
  scale_fill_discrete(name="Education") +
  guides(color="none") +
  ggtitle("Perceived sensitivity of mental/affective information by socioeconomical status") +
  theme_cowplot() +
  theme(plot.margin = margin(5, 10, 5, 5)) +
  NULL

ggsave("figures/SI_SES_mental_affective.tiff", dpi=600, width=10.5, height=5)

```


## - does how early one adopts tech predicting perceived sensitivity (factors) 
```{r}
contrasts(df_sens_fa_long$Factor_name) <- contr.sum(3)
m_fs_by_ear <- lmer(Factor_score_wa ~ 
                      Factor_name * scale(EarAdoptLvl) + (1|ID),
                    data=df_sens_fa_long)

### anova result - only main effect of factor (already known)
Anova(m_fs_by_ear, type=3, test.statistic="F")
# summary(m_fs_by_ear) %>% coef() %>% round(4)

### post-hoc comparisons
emm_options(pbkrtest.limit = 5000)
emm_options(lmerTest.limit = 5000)
et <- emtrends(m_fs_by_ear, ~ Factor_name, var = "EarAdoptLvl")
test(et)
confint(m_fs_by_ear, "beta_",
        parm="scale(EarAdoptLvl)",
        level=0.95, method="boot", nsim=1000, boot.type="perc",
        FUN=NULL, quiet=FALSE, oldNames=TRUE) %>%
  round(2)

```


## - plot
```{r}
df_plot_m_fs_by_ear <- ggemmeans(m_fs_by_ear, EarAdoptLvl ~ Factor_name) %>% 
  data.frame() %>% 
  mutate(group=factor(group, 
                      levels=c("Private", 
                               "Mental/Affective", 
                               "Public"))) %>%
  rename(Factor_name=group)

df_plot_m_fs_by_ear %>%
  ggplot(aes(x=x, y=predicted, fill=Factor_name, color=Factor_name)) +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high, color=NULL), alpha=0.3) +
  geom_line() +
  scale_x_continuous(name="How early do you adopt new technologies relative to others?",
                     breaks=1:5,
                     labels=c("1=first",
                              "2=among the first",
                              "3=middle",
                              "4=among the last",
                              "5=last")) +
  scale_y_continuous(name="Perceived sensitivity", 
                     limits = c(1, 4),
                     breaks = 1:4,
                     labels = c("1=\nnot at all sensitive",
                                "2=\nnot too sensitive",
                                "3=\nsomewhat sensitive",
                                "4=\nvery sensitive")) +
  scale_color_manual(values=c(
    "Public"="#00BA38",
    "Private"="#619CFF",
    "Mental/Affective"="#F8766D")) +
  scale_fill_manual(name="Type of Information",
                    values=c(
                      "Public"="#00BA38",
                      "Private"="#619CFF",
                      "Mental/Affective"="#F8766D")) +
  guides(color="none") +
  ggtitle("Early tech adoption had no correlation with perceived sensitivity of information") +
  theme_cowplot() +
  NULL

ggsave("figures/Factor scores by Early Tech Adoption.tiff",
       width = 11, height = 6, dpi = 600)

```


